---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Disable scientific notation
options(scipen = 999)

# Avoid conflicts
summarize <- dplyr::summarize
rename <- dplyr::rename
count <- dplyr::count

# Define the base path to your CSV files
df <- "/Users/gabrielmancillas/Desktop/ADS 505-01/FINAL/Police_settlements/Locations/selected_data.csv"
```

```{r}
#load the library
library(readr)

# Load the data from the CSV file
data <- read_csv(df)

# Display the first few rows of the data to verify and column names
head(data)

```

```{r}

# Install dplyr if not already installed
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}

# Load dplyr for piping functionality
library(dplyr)
```
```{r}
# Define a function to check for these special keywords and replace the summary_allegations
replace_with_keyword <- function(text) {
  if (grepl("civil rights", text, ignore.case = TRUE)) {
    return("Civil Rights")
  } else if (grepl("declaratory judgment", text, ignore.case = TRUE)) {
    return("Declaratory Judgment")
  } else if (grepl("litigation", text, ignore.case = TRUE)) {
    return("Litigation")
  } else if (grepl("personal injury", text, ignore.case = TRUE)) {
    return("Personal Injury")
  } else if (grepl("negligence", text, ignore.case = TRUE)) {
    return("Negligence")
  } else if (grepl("presuit", text, ignore.case = TRUE)) {
    return("Presuit")
  } else if (grepl("brutality", text, ignore.case = TRUE)) {
    return("Brutality")
  } else if (grepl("cruel & unusual", text, ignore.case = TRUE)) {
    return("Cruel & Unusual")
  } else if (grepl("due process", text, ignore.case = TRUE)) {
    return("Due Process")
  } else if (grepl("errors & omissions", text, ignore.case = TRUE)) {
    return("Errors & Omissions")
  } else if (grepl("fail to render care", text, ignore.case = TRUE)) {
    return("Fail to Render Care")
  } else if (grepl("failure to enforce", text, ignore.case = TRUE)) {
    return("Failure to Enforce")
  } else if (grepl("false arrest", text, ignore.case = TRUE)) {
    return("False Arrest")
  } else if (grepl("improper auto procedure", text, ignore.case = TRUE)) {
    return("Improper Auto Procedure")
  } else if (grepl("miscellaneous", text, ignore.case = TRUE)) {
    return("Miscellaneous")
  } else if (grepl("negligent supervision", text, ignore.case = TRUE)) {
    return("Negligent Supervision")
  } else if (grepl("search & seizure", text, ignore.case = TRUE)) {
    return("Search & Seizure")
  } else {
    return(text)  # If no keyword is found, return the original text
  }
}

# Apply the function to the summary_allegations column
data <- data %>%
  mutate(summary_allegations = sapply(summary_allegations, replace_with_keyword))
#show the data
print(data)
```


# Define the replace_with_keyword function
replace_with_keyword <- function(text) {
  if (grepl("excessive force", text, ignore.case = TRUE)) {
    return("Excessive Force")
  } else if (grepl("unlawful arrest", text, ignore.case = TRUE)) {
    return("Unlawful Arrest")
  } else if (grepl("unlawful entry", text, ignore.case = TRUE)) {
    return("Unlawful Entry")
  } else if (grepl("unlawful seizure", text, ignore.case = TRUE)) {
    return("Unlawful Seizure")
  } else if (grepl("malicious prosecution", text, ignore.case = TRUE)) {
    return("Malicious Prosecution")
  } else if (grepl("42 U.S.C.|1983 case", text, ignore.case = TRUE)) {
    return("Civil Rights Violation")
  } else if (grepl("4th Amendment Violations", text, ignore.case = TRUE)) {
    return("4th Amendment Violations")
  } else {
    return(text)  # Return the original text if no match
  }
}

# Apply the function to rows 1 to 42
data <- data %>%
  slice(1:42) %>%  # Select rows 1 to 42
  mutate(summary_allegations = sapply(summary_allegations, replace_with_keyword))

# If you want to apply the transformation only to rows 1-42 and leave the rest unchanged:
combined_data <- data %>%
   mutate(summary_allegations = ifelse(row_number() <= 42, 
                                       sapply(summary_allegations, replace_with_keyword), 
                                       summary_allegations))

# View the transformed data
table(head(combined_data$summary_allegations, 42))  # View the first 42 rows
```
```{r}
library(dplyr)

# Define a function to standardize summary allegations
replace_with_keyword <- function(text) {
  if (grepl("civil rights", text, ignore.case = TRUE)) {
    return("Civil Rights")
  } else if (grepl("wrongful arrest", text, ignore.case = TRUE)) {
    return("Wrongful Arrest")
  } else if (grepl("negligence", text, ignore.case = TRUE)) {
    return("Negligence")
  } else if (grepl("police shooting", text, ignore.case = TRUE)) {
    return("Police Shooting")
  } else if (grepl("wrongful demolition", text, ignore.case = TRUE)) {
    return("Wrongful Demolition")
  } else if (grepl("other", text, ignore.case = TRUE)) {
    return("Other")
  } else {
    return(text)  # Return original if no match is found
  }
}

# Apply the function to the summary_allegations column
data <- data %>%
  mutate(summary_allegations = sapply(summary_allegations, replace_with_keyword))

# Group by the new, standardized summary_allegations and count occurrences
grouped_data <- data %>%
  group_by(summary_allegations) %>%
  summarise(count = n())

# View the grouped data with counts
print(grouped_data)
# show the entire data
print(data)
head(data)


____________________________________________________________
# Step 2: Define the function to standardize summary allegations
replace_with_keyword <- function(text) {
  if (grepl("civil rights|equity|constituional claim|constitutional violations|CIVIL RIGHTS CLAIMS|Civil Rights Violation", text, ignore.case = TRUE)) {
    return("Civil Rights")
  } else if (grepl("wrongful arrest|prisoner complaint common law", text, ignore.case = TRUE)) {
    return("Wrongful Arrest")
  } else if (grepl("Violation of constitutional rights:", text, ignore.case = TRUE)) {
    return("Violation of Constitutional Rights")
  } else if (grepl("Uncoded", text, ignore.case = TRUE)) {
    return("Uncoded")
  } else if (grepl("excessive force", text, ignore.case = TRUE)) {
    return("Excessive Force")
  } else if (grepl("search & seizure", text, ignore.case = TRUE) | grepl("unlawful search", text, ignore.case = TRUE)) {
    return("Search & Seizure")
  } else if (grepl("false imprisonment|False Arrest/Detention/Imprisonment", text, ignore.case = TRUE)) {
    return("False Imprisonment")
  } else if (grepl("assault and battery|Assault/Slander/Liable|property damage|Excessive Force (Police)|Law Enforcement Related Injury - Assault & Battery/Excessive Force", text, ignore.case = TRUE)) {
    return("Assault & Battery")
  } else if (grepl("malicious prosecution", text, ignore.case = TRUE)) {
    return("Malicious Prosecution")
  } else if (grepl("negligence", text, ignore.case = TRUE)) {
    return("Negligence")
  } else if (grepl("Disability/Medical Discrimination", text, ignore.case = TRUE)) {
    return("Disability/Medical Discrimination")
  } else if (grepl("police shooting|Wrongful Death", text, ignore.case = TRUE)) {
    return("Police Shooting")
  } else if (grepl("unlawful detention", text, ignore.case = TRUE)) {
    return("Unlawful Detention")
  } else if (grepl("police misconduct|officer misconduct|police activity|Police False Arrest-", text, ignore.case = TRUE)) {
    return("Police/Officer Misconduct")
  } else if (grepl("motor Vehicle Incident|Struck by Police Cruiser|car accident|Breathalyzer|auto", text, ignore.case = TRUE)) {
    return("Motor Vehicle Incident")
  } else if (grepl("failure to enforce", text, ignore.case = TRUE)) {
    return("Failure to Enforce")
} else if (grepl("errors & omissions", text, ignore.case = TRUE)) {
    return("Errors & Omissions")
  } else if (grepl("malicious prosecution", text, ignore.case = TRUE)) {
    return("Malicious Prosecution")
  } else if (grepl("negligent|failure to provide medical care|fail to render medcare|failure to prevent attempts of suicide", text, ignore.case = TRUE)) {
    return("Negligence")
  } else if (grepl("miscellaneous|trash container|whistle blower action", text, ignore.case = TRUE)) {
    return("Miscellaneous")
  } else if (grepl("improper auto procedure|improper procedure|impropr procedure", text, ignore.case = TRUE)) {
    return("Improper Auto Procedure")
  } else if (grepl("cruel", text, ignore.case = TRUE)) {
    return("Wrongful Arrest")
  } else if (grepl("brutality|Personal Injury: Exessive force|T.R.O.", text, ignore.case = TRUE)) {
    return("Excessive Force")
  } else if (grepl("false arrest|false imprisonment", text, ignore.case = TRUE)) {
    return("False Imprisonment")
  } else if (grepl("due process", text, ignore.case = TRUE)) {
    return("Other")
  } else if (grepl("DUI stop", text, ignore.case = TRUE)) {
    return("Civil Rights")
  } else if (grepl("illegal search|seizure", text, ignore.case = TRUE)) {
    return("Search & Seizure")
  } else if (grepl("reversed conviction|Unjust Imprisonment Act claim", text, ignore.case = TRUE)) {
    return("Wrongful Arrest")
  } else if (grepl("litigation|presuit|Subrogation claim", text, ignore.case = TRUE)) {
    return("Litigation")  
  } else if (grepl("sexual harrasement|sexual assault", text, ignore.case = TRUE)) {
    return("Sexual Harassment")
  } else if (grepl("personal injury|PEACE OFFICER/POLICE ACTION (PI)", text, ignore.case = TRUE)) {
    return("Personal Injury")
  } else if (grepl("declaratory judgment", text, ignore.case = TRUE)) {
    return("Declaratory Judgment")
  } else if (grepl("Employment Complaint-Common Law|(non-labor/employment)", text, ignore.case = TRUE)) {
    return("Employment Complaint")
  } else if (grepl("Violation of constitutional rights:|Violation of constitutional rights:|Amendments", text, ignore.case = TRUE)) {
    return("Violation of constitutional rights")
  } else if (grepl("Discrimination - sex", text, ignore.case = TRUE)) {
    return("Race Discrimination")
  } else if (is.na(text)) {
    return("Not Reported")# Return the original text if no match is found
  } else {
    return(text)
  }
}
